---
title: "p8105_hw3_gl2761"
author: "Gonghao Liu"
date: "10/15/2022"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

### Problem 2

```{r}
accel_df = read_csv("accel_data.csv")

accel_clean = accel_df %>% 
  janitor::clean_names() %>% 
  mutate(
    weekday_weekend = case_when(
    day == "Monday" ~ "Weekday",
    day == "Tuesday" ~ "Weekday",
    day == "Wednesday" ~ "Weekday",
    day == "Thursday" ~ "Weekday",
    day == "Friday" ~ "Weekday",
    day == "Saturday" ~ "Weekend",
    day == "Sunday" ~ "Weekend",
  ),
  day = as.factor(day),
  day = fct_relevel(day,
                    c("Monday", "Tuesday", "Wednesday", "Thursday", 
                      "Friday", "Saturday", "Sunday"))) %>% 
  relocate(week, day_id, day, weekday_weekend) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_min",
    names_prefix = "activity_",
    values_to = "activity_counts"
  ) %>% 
  mutate(
    activity_min = as.numeric(activity_min),
    weekday_weekend = as.factor(weekday_weekend)
    )
  
```

Description of resulting dataset:
This data set has `r ncol(accel_clean)` variables and `r nrow(accel_clean)` observations. 

Variables are `r names(accel_clean)`.

Here are variable descriptions:

`r colnames(accel_clean)[1]`: Number of week.

`r colnames(accel_clean)[2]`: Number of day.

`r colnames(accel_clean)[3]`: the day that related to `r colnames(accel_clean)[2]`.

`r colnames(accel_clean)[4]`: Whether it is a weekday or weekend.

`r colnames(accel_clean)[5]`: Each activity minute.

`r colnames(accel_clean)[6]`: The counts on each activity minute.

```{r}
accel_clean %>% 
  group_by(week, day) %>% 
  summarise(total_activity = sum(activity_counts),
            .groups = "drop") %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity
  ) %>% 
  knitr::kable(caption = "Total activity data over the day") 
```

I don't think there seem to be any trends in the existing data.

```{r}
accel_clean %>% 
  ggplot(aes(x = activity_min,y = activity_counts,color = day)) +
  geom_point(alpha = .5) +
  labs(
    title = "24_hour activity time courses for eah day",
    x = "Time of the day",
    y = "Activity time counted"
  ) +
  viridis::scale_color_viridis(
    name = "Day",
    discrete = TRUE) +
    scale_x_continuous(
    breaks = c(0, 120, 240, 360, 480, 600, 720, 840, 960, 1080, 1200, 1320, 1440),
    labels = c("0h", "2h", "4h", "6h", "8h", "10h", "12h", "14h", 
               "16h", "18h", "20h", "22h", "24h")
  ) +
  theme(plot.title = element_text(hjust = .5))
```

Based on the graph above, we can see that this patient starts to be active around 7 am, around noon, around 4:30 pm on Thursday, Saturday and Sunday, and is active around 9 pm on most days except Sunday. 

### Problem 3

```{r}
data("ny_noaa")
skimr::skim(ny_noaa)
```

Clean the data 
```{r}
ny_noaa_clean = 
  ny_noaa %>%
  janitor::clean_names()%>%
  separate(date, into = c("year", "month", "day"))%>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.name[month],
    day = as.integer(day),
    tmax = as.numeric(tmax)/10,
    prcp = as.numeric(prcp)/10,
    tmin = as.numeric(tmin)/10
    )%>%
  relocate(year, month, day) 
```

```{r}
ny_noaa_clean %>%
  count(snow) %>%
  arrange(desc(n))
```
The most commonly observed value for the snow variable is 0 because most of the weather stations didn't observe snow most of the time.

```{r}
ny_noaa_clean %>% 
  filter(month %in% c('January', 'July')) %>%
  filter(!is.na(tmax)) %>% 
  group_by(month, year, id) %>% 
  summarise(mean_temp = mean(tmax)) %>% 
  ggplot(aes(x = as.factor(year), y=mean_temp, color = year)) +
  geom_point() +
  geom_density(alpha = .5) +
  facet_grid(.~ month) +
  labs(
    title = "Distribution of average max temperature for January and July for each station",
    x = "Years",
    y = "Average max temperature",
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 5))
```

The average temperature in July is higher than in January. In 1982 there is a day in January that is extremely cold day. In 1988 there is a day much cooler than usual.

```{r}
ny_noaa_clean%>%
  filter(!is.na(tmax)| !is.na(tmin)) %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex() +
  labs(
    title = "tmax vs tmin for whole dataset",
    x = "Max Temperature",
    y = "Min Temperature",
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 5))
```

Tmax vs Tmin has the most density in the region of (0,0) and (15,25).

```{r}
ny_noaa_clean%>%
  filter(!is.na(snow)) %>%
  filter(snow>0 & snow<100)%>%
  group_by(year)%>%
  ggplot(aes(x=snow, color=as.factor(year))) +
  geom_density(alpha=0.5)+
  labs(
    title = "Distribution of snow by year",
    x = "Snowfall(mm)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 5))
```

According to the graph above, the snowfall density has a similar trend in all the years. Snowfall gets to it's largest density around 12mm and 25mm.